---
title: "Gap analysis"
author: "Ivan Gonzalez"
date: "Wednesday, December 02, 2015"
output: html_document
---

This is an code adaptation to develop an spatial gap index analysis sensu [*A methodological framework to quantify the spatial quality of biological databases*](http://www.biodiversity-plants.de/biodivers_ecol/article_meta.php?DOI=10.7809/b-e.00057)

The follow R code is an adaptation to get a final gap index map.

The first step is load information. For this is required to have both a database with biological records and the GIS layers.
The GIS information can be retived from [climatic folder](https://drive.google.com/open?id=0B3J9l9VrPJBRTzV0Z0w5ZUd0bVE) and [bias factors](https://drive.google.com/open?id=0B3J9l9VrPJBRSjg1b3BEZDhETUE). 
The datase must be ´data.frame´ class table with the follow columns:
```{r, echo = TRUE}
setwd('C:/IAvH/VACIOS')
```


```{r, cache = TRUE}
DATOS <- read.csv('root/DATOS.csv', as.is = TRUE)
head(DATOS)
```

For replicate this excercice consider change the folder paths, check the encoding format and separator values and install libraries dependences.

The initial step is prepare data, paths and functions and load libraries.
```{r, cache = TRUE}
# ---------------------------
# 0.  Data loading
# ---------------------------
library(raster)
library(maptools)
library(rgdal)
library(SPECIES)
library(classInt)

source('root/GAPfunctions.R')
outPlotDir <- 'root/plots'
ruta_factores <- 'root/GIS/FACTORES'
ruta_ambientales <- 'root/GIS/AMBIENTALES'

DATOS2 <- DATOS[, c('ID', 'lat', 'lon')]
coordinates(DATOS2)=~lon+lat

colombia <- readShapePoly(paste0(ruta_factores, "/COLOMBIA.shp"))
grilla <- raster(paste0(ruta_factores, "/grilla 10k.img"))
grilla[] <- 1:ncell(grilla)

en_area <- mask(grilla, colombia)

## Assign the pixel ID to each coordinate
celdas <- extract(en_area, DATOS2)
DATOS$celdas <- celdas

rm(celdas)
```


As the article show, ther's tree components to calculet the gap index. The first is the record density and is estimated as the number of records over each pixel:
```{r, cache = TRUE}
# ---------------------------
# 1.  Records density 
# ---------------------------

DATOS2$CONT <- 1
densRegistros <- rasterize(DATOS2, en_area, field = DATOS2$CONT, fun = sum, na.rm = TRUE)
densRegistros[is.na(densRegistros[])] <- 0
densRegistros <- mask(densRegistros, colombia, "INDICE_DENSIDAD.tif", overwrite = TRUE)

# Let's generate a png file with the density plot
data(wrld_simpl)
par(mfrow=c(1,1))

png(file = paste0(outPlotDir, "/1-INDICE_densidad.png")) 
plot(wrld_simpl, main="UBICACION DE LOS REGISTROS \n BASE DE DATOS INICIAL", col = 'lightyellow')
points(x=coordinates(DATOS2)[, 1], y=coordinates(DATOS2)[, 2], col = rgb(139, 0, 0, 100, maxColorValue = 255), cex = .1, pch = 20)

plot(colombia, main = "UBICACION DE LOS REGISTROS COLOMBIA")
plot(DATOS2, add = TRUE, pch = 20)
dev.off()

# Let's generate a png file with the density plot. Could take several minutes becouse of table size
pdf(file = paste0(outPlotDir, "/1.INDICE_densidad.pdf"))
plot(densRegistros, col = rev(topo.colors(10)), main="DENSIDAD DE  PUNTOS/ 10Km2")
plot(colombia, add = TRUE)
density(densRegistros, plot = TRUE)
persp(densRegistros, xlab = "X coordinates", ylab = "Y coordinates", zlab = "density",
      phi = 35, theta = 20, col = "lightblue", expand = .5, ticktype = "detailed")
dev.off()

plot(densRegistros, col = rev(topo.colors(10)), main = "DENSIDAD DE  PUNTOS/ 10Km2")
plot(colombia, add = TRUE)

rm(wrld_simpl)

# Keep 'densRegistros' object
```

The second element for the gap index is a bias layer

```{r, cache = TRUE}
# ---------------------------
# 2.  Bias in sampling  
# ---------------------------
DATOS_s <- unique(DATOS[, c('lat', 'lon')])
N <- nrow(DATOS_s)

# ---------------------------
# 2.2. Bias by phisical factors
# ---------------------------

AP <- readOGR(ruta_factores, "protectedAreas")
urbano <- readOGR(ruta_factores, "urban")
rios <- readOGR(ruta_factores, "riversMain")
vias <- readOGR(ruta_factores, "roads")

pdf(file = paste0(outPlotDir, "/2.1 INDICE_factores_sesgos.pdf "))
par(mfrow = c(1, 2))
sesgo_ap <- BIAS(biasLayer = AP, rasterMask = grilla, layerName = "AREA_PROTEGIDA",  outDir = outPlotDir)
sesgo_rios <- BIAS(biasLayer = rios, rasterMask = grilla, layerName = "RIOS",  outDir = outPlotDir)
sesgo_vias <- BIAS(biasLayer = vias, rasterMask = grilla,layerName = 'VIAS', outDir = outPlotDir)
sesgo_urbano <- BIAS(biasLayer = urbano, rasterMask = grilla, layerName = "CASCOS_URBANOS", outDir = outPlotDir)
dev.off()

# Remove extra files and keep sesgo_ap, sesgo_rios, sesgo_vias, sesgo_urbano
rm(AP, urbano, rios, vias, DATOS_s, sesgo_ap, sesgo_rios, sesgo_vias, sesgo_urbano, grilla)

# ---------------------------
# 2.2. Bias by environment
# ---------------------------

# 2.2.1 Compare distributions
## Compare environment all distribution values and sampled values by kolmogorov and Kullback-Leibler divergence

ambientales <- stack(paste0(ruta_ambientales, '/', c(paste0("bio_", 1:19), "alt", "slope")))
envVarNames <- c("TEMPERATURA MEDIA ANUAL", "MEDIA RANGO DIURNO", "ISOTERMALIDAD", "ESTACIONALIDAD TEMPERATURA",
             "MAX T MES MAS CALIDO", "MIN T MES FRIO", "RANGO ANUAL T", "T MEDIA  DEL CUARTO HUMEDO", 
             "T MEDIA DEL CUARTO SECO", "T MEDIA DEL CUARTO CALIDO", "T M DEL CUARTO FRIO", "PRECIPITACI?N ANUAL", 
             "PP MES MAS HUMEDO", "PP MES MAS SECO", "PP ESTACIONAL", "PP CUARTO HUMEDO", "PP CUARTO SECO",
             "PP CUARTO CALIDO", "PP CUARTO FRIO", "ALTURA", "PENDIENTE")

# Generate data.frame with all values
VARIABLES <- ambientales[1:ncell(ambientales)]
VARIABLES <- na.omit(VARIABLES) 

# Generate data.frame with observed values
MUESTREO <- extract(ambientales, DATOS2)
MUESTREO <- na.omit(MUESTREO)

pdf(file = paste0(outPlotDir, "/2.2 INDICE_COMPARACION_VARIABLES.pdf ")) 
par(mfrow = c(4, 3))

numPredictors <- dim(ambientales)[3]
RESULT <- NULL

for (i in 1:numPredictors){
  
  # Generate a random vector of complete variable with sampled vector size
  sample.Var.i <- sample(VARIABLES[, i], nrow(MUESTREO), replace = T)

  # Compare both random and sampled vector
  COMPARACION <- ks.test(sample.Var.i, MUESTREO[, i]) 

  VAR <- envVarNames[i]
  media_var <- mean(VARIABLES[, i])
  media_DATOS <- mean(MUESTREO[, i])
  des_var <- sd(VARIABLES[, i])
  des_DATOS <- sd(MUESTREO[, i])
  coefvar_var <- des_var/media_var
  coefvar_DATOS <- des_DATOS/media_DATOS
  D <- COMPARACION$statistic
  Pval <- COMPARACION$p.value
  if (Pval <= 0.05){ 
    DESICION <- "Distribuciones diferentes"
  } else {
    DESICION <- "Distribuciones iguales"
  }
  Nceldas <- ncell(VARIABLES[, i])
  
  PI <- as.data.frame(table(MUESTREO[, i])/nrow(MUESTREO))  
  QTOTAL <- as.data.frame(table(VARIABLES[, i])/Nceldas)
  QI <- QTOTAL[match(PI[, 1], QTOTAL[, 1]), ]
  
  Q <- cbind(QI[, 2], PI[, 2])
  Q <- as.matrix(Q)
  #KLdiv(Q,overlap=F, method="discrete")
  #DIVERGENCIA=KLdiv(Q,overlap=F,method="discrete",na.rm=T)[1,2]
  
  DENSI_VAR <- density(VARIABLES[, i])
  DENSI_MUES <- density(MUESTREO[, i])
  minx <- min(min(DENSI_VAR$x), min(DENSI_MUES$x))
  maxx <- max(max(DENSI_VAR$x), max(DENSI_MUES$x))
  miny <- min(min(DENSI_VAR$y), min(DENSI_MUES$y))
  maxy <- max(max(DENSI_VAR$y), max(DENSI_MUES$y))
  
  plot(density(VARIABLES[,i]), col='black', xlab = '', main=VAR, xlim=c(minx, maxx), ylim=c(miny,maxy),
       sub = bquote(p(H[a]):  ~ .(Pval)))  
  lines(density(MUESTREO[,i]), col= "red", main=VAR)
  RESUMEN <- cbind(VAR, media_var, media_DATOS, des_var, des_DATOS, coefvar_var, coefvar_DATOS,
                   D, Pval, DESICION)
  RESULT <- rbind(RESULT, RESUMEN)
}

# Write table with summary for each variable
write.table(RESULT, "bondad ajuste variables.txt", sep="\t",col.names=T)

# Remove extra objects
rm(MUESTREO, VARIABLES, RESULT, RESUMEN, VAR, media_var, media_DATOS, des_var, des_DATOS, coefvar_var, coefvar_DATOS,
   D, Pval, DESICION, DENSI_VAR, DENSI_MUES, sample.Var.i, Q, QI, QTOTAL)

dev.off()

# 2.2.2. Environmental bias layer

pdf(file = paste0(outPlotDir, "/2.3 INDICE_sesgos_ambientales.pdf ")) ##comienza la grafica tipo pdf
par(mfrow=c(1,2))
d <- ambientales[[1]] * 0
resultados <- NULL

localidades <- rasterize(DATOS2, en_area, fun = function(x, ...) {length(unique(na.omit(x)))})
recla <- as.matrix(cbind(1, localidades@data@max, 1))
localidades <- reclassify(localidades, recla)
sesgosDF <- NULL
for (k in 1:numPredictors){
  ses_amb <- BIAS(biasLayer = ambientales[[k]], layerName = names(ambientales)[k], outDir = outPlotDir, doplot = TRUE)
  d <- sum(d, ses_amb$biasLayer, na.rm = TRUE)
  sesgosDF <- cbind(sesgosDF, ses_amb$biasLayer[])
  resultados <- cbind(resultados, ses_amb$sesgos)
}
colnames(sesgosDF) <- names(resultados) <- envVarNames

dev.off()

pcaLayer <- raster(paste0(ruta_ambientales, '/pcaLayer.tif'))
ses_pca <- BIAS(biasLayer = pcaLayer, layerName = 'pca', outDir = outPlotDir, doplot = TRUE)
d2 <- ses_pca$biasLayer

write.table(resultados, "sesgos_ambientales.txt", sep="\t", col.names = TRUE)
write.table(sesgosDF, "sesgos_tabla_Capas.txt", sep="\t", col.names = TRUE)
writeRaster(d, "INDICE_AMBIENTAL.tif", overwrite = TRUE)

```

The final component is the species completness. For this case we are gonna to use Jackknife and bootstrap estimates for two databse columns: original species and accepted species name.

```{r, cache = TRUE}
# ---------------------------
# 3. Data base completness
# ---------------------------

spListByCell <- DATOS[!is.na(DATOS$celdas), c('species', 'celdas')]
spListByCellHQ <- DATOS[!is.na(DATOS$celdas) & !is.na(DATOS$acceptedNameUsage), c('acceptedNameUsage', 'celdas')]

freqTable <- table(spListByCell$celdas)
freqTableHQ <- table(spListByCellHQ$celdas)

treshold <- 0

spListByCell <- spListByCell[spListByCell$celdas %in% names(which(freqTable >= treshold)), ]
spListByCellHQ <- spListByCellHQ[spListByCellHQ$celdas %in% names(which(freqTableHQ >= treshold)), ]

estimateS <- richEst(spListByCell$species, spListByCell$celdas)
estimateSHQ <- richEst(spListByCellHQ$acceptedNameUsage, spListByCellHQ$celdas)

rm(spListByCell, spListByCellHQ)

compRichBoot <- compRichJack <- richJackHQ <- richBootHQ <- richJack <- richBoot <- en_area * 0
richBoot[as.numeric(rownames(estimateS))] <- estimateS$Boot
richJack[as.numeric(rownames(estimateS))] <- estimateS$JNhat
richBootHQ[as.numeric(rownames(estimateSHQ))] <- estimateSHQ$Boot
richJackHQ[as.numeric(rownames(estimateSHQ))] <- estimateSHQ$JNhat

compRichBoot[as.numeric(rownames(estimateS))] <- estimateS$Sobs/estimateS$Boot
compRichJack[as.numeric(rownames(estimateS))] <- estimateS$Sobs/estimateS$JNhat

#compRichBoot[as.numeric(rownames(estimateS)[estimateS$Sobs == 1])] <- NA
#compRichJack[as.numeric(rownames(estimateS)[estimateS$Sobs == 1])] <- NA
compRichJack[compRichJack[] >= 1] <- 1
compRichBoot[compRichBoot[] >= 1] <- 1

pdf(file = paste0(outPlotDir, "/3 INDICE_COMPLEMENTARIEDAD.pdf ")) 
par(mfrow=c(2, 2)) 

plot(richBoot, main = paste('Bootstrap:', nrow(estimateS)))
plot(richJack, main = paste('Jackknife:', nrow(estimateS)))

plot(richBootHQ, main = paste('BootstrapHQ:', nrow(estimateSHQ)))
plot(richJackHQ, main = paste('JackknifeHQ:', nrow(estimateSHQ)))


plot(compRichBoot, main = paste('Bootstrap:', nrow(estimateS)))
plot(colombia, add = TRUE)
plot(compRichJack, main = paste('Jackknife:', nrow(estimateS)))
plot(colombia, add = TRUE)

richBootVals <- compRichBoot[!is.na(compRichBoot[]) & compRichBoot[] != 0]
richJackVals <- compRichJack[!is.na(compRichJack[]) & compRichJack[] != 0]
hist(richBootVals, main = 'Density Bootstrap', freq = FALSE, xlim = c(0, 1.2))
lines(density(richBootVals), main = 'Density Bootstrap')

hist(richJackVals, main = 'Density Jackknife', freq = FALSE, xlim = c(0, 1.2))
lines(density(richJackVals), main = 'Density Jackknife')
dev.off()

writeRaster(compRichBoot, "INDICE_COMPLEMENTARIEDAD_JACK.tif", overwrite=TRUE)
writeRaster(compRichJack, "INDICE_COMPLEMENTARIEDAD_BOOTS.tif", overwrite=TRUE)
```


Finally sum the three standardized components using the article formula

```{r, cache = TRUE}
# ---------------------------
# 4. GSI / Gap selection index
# ---------------------------

# Layer standardization

DENSIDAD <- normalize01(densRegistros)
AMBIENTAL <- normalize01(d)
COMPLEMENTARIEDAD_BOOT <- normalize01(compRichBoot)
COMPLEMENTARIEDAD_JACK <- normalize01(compRichJack)

writeRaster(DENSIDAD,"INDICE_DENSIDAD_EST.tif", overwrite=TRUE)
writeRaster(AMBIENTAL,"INDICE_AMBIENTAL_EST.tif", overwrite=TRUE)
writeRaster(COMPLEMENTARIEDAD_BOOT,"INDICE_COMPLEMENTARIEDAD_BOOT_EST.tif", overwrite=TRUE)
writeRaster(COMPLEMENTARIEDAD_JACK,"INDICE_COMPLEMENTARIEDAD_JACK_EST.tif", overwrite=TRUE)

# Plot 
## Records ubication
png(file = paste0(outPlotDir, "/4 INDICE_final_ubicacion_registros.png")) ##comienza la grafica tipo pdf 
par(mfrow = c(1, 1))
plot(colombia, main = "UBICACION DE \n LOS REGISTROS")
plot(DATOS2, add = T, pch = 20, cex = 0.1, col = rgb(0.1, 0.2, 1, 0.01))
dev.off()

pdf(file = paste0(outPlotDir, "/4 INDICE_final.pdf ")) ##comienza la grafica tipo pdf 

## Records density
par(mfrow=c(1, 2))
plot(densRegistros, main = 'Records density')
plot(colombia, add = TRUE)
plot(DENSIDAD, main = 'Records density')
plot(colombia, add = TRUE)

## Environmental bias
par(mfrow=c(1, 2))
plot(mask(d, colombia), main = 'd Index')
plot(colombia, add = TRUE)
plot(mask(AMBIENTAL, colombia), main = 'd Index \n Normalized')
plot(colombia, add = TRUE)

par(mfrow=c(2, 2))
dValsNoNA <- d[!is.na(d[])]
dValsNoNANoZeros <- d[!is.na(d[]) & d[] != 0]
hist(dValsNoNA, main = 'd Index no NA no 0')
hist(dValsNoNANoZeros, main = 'd Index \n no NA no 0')

AMBValsNoNA <- AMBIENTAL[!is.na(AMBIENTAL[])]
AMBValsNoNANoZeros <- AMBIENTAL[!is.na(AMBIENTAL[]) & AMBIENTAL[] != 0]
hist(AMBValsNoNA, main = 'Ambiental no NA')
hist(AMBValsNoNANoZeros, main = 'Ambiental\nno NA no 0')

## Completness
par(mfrow = c(1, 2))
plot(compRichBoot, main = 'Bootstrap')
plot(colombia, add = TRUE)
plot(compRichJack, main = 'Jackknife')
plot(colombia, add = TRUE)

par(mfrow=c(2, 2))

bootValsNoNA <- compRichBoot[!is.na(compRichBoot[])]
bootValsNoNANoZeros <- compRichBoot[!is.na(compRichBoot[]) & compRichBoot[] != 0]
hist(bootValsNoNA, main = 'Bootstrap no NA')
hist(bootValsNoNANoZeros, main = 'Bootstrap\nno NA no 0')

JacValsNoNA <- compRichJack[!is.na(compRichJack[])]
JacValsNoNANoZeros <- compRichJack[!is.na(compRichJack[]) & compRichJack[] != 0]
hist(JacValsNoNA, main = 'Jackknife no NA no 0')
hist(JacValsNoNANoZeros, main = 'Jackknife no NA no 0')


#  GSI / GAP INDEX

GSI_BOOT <- (3 - DENSIDAD - AMBIENTAL - COMPLEMENTARIEDAD_BOOT)/3
GSI_JACK <- (3 - DENSIDAD - AMBIENTAL - COMPLEMENTARIEDAD_JACK)/3

par(mfrow = c(1, 2))

plot(GSI_BOOT, main = 'GAP SELCTION INDEX (BOOT)', zlim = c(0, 1))
plot(colombia, add = TRUE)

plot(GSI_JACK, main = 'GAP SELCTION INDEX (JACK)', zlim = c(0, 1))
plot(colombia, add = TRUE)

densJack <- density(GSI_JACK, main = 'Jackk', xlim = c(0, 1))
densBoot <- density(GSI_BOOT, main = 'Boot', xlim = c(0, 1))

par(mfrow = c(1, 1))
plot(densJack, main = 'GSI values', col = 'blue', ylim = c(0, 10))
lines(densBoot, col = 'red')
legend('topleft', legend = c('Jackknife', 'Bootstrap'), 
       lty = c(1, 1), lwd = c(1, 1), col = c('blue', 2))

dev.off()

writeRaster(GSI_BOOT,"INDICE_GSI_BOOT.tif", overwrite = TRUE)
writeRaster(GSI_JACK,"INDICE_GSI_JACK.tif", overwrite = TRUE)

```
